### 2022-02-09 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@e56623fbe278bf9e6b31b7a9183dc8d59cf0db2f 🚀

Being able to infer \emph{potential} interactions could be the catalyst
for significant breakthroughs in our ability to start thinking about
species interaction networks over large spatial scales (Hortal et al.,
2015). In a recent overview of the field of ecological network
prediction, Strydom, Catchen, et al. (2021) identified two challenges of
interest to the prediction of interactions at large scales. First, there
is a relative scarcity of relevant data in most places globally --
paradoxically, this restricts our ability to infer interactions to
locations where inference is perhaps the least required; second,
accurate predictions often demand accurate predictors, and the lack of
methods that can leverage small amount of data is a serious impediment
to our global predictive ability. In most places, our most reliable
biodiversity knowledge is that of a species pool: through the analysis
of databases like GBIF or IUCN, it is possible to establish a list of
species in a region of interest; but establishing the interactions
between these species is difficult.

@@

\begin{figure}
\hypertarget{fig:embedding}{%
\centering
\includegraphics{figures/conceptual_embedding.png}
\caption{Overview of the embedding process. A network (\emph{A}),
possibly represented as its adjacency matrix (\emph{B}), is converted
into a lower-dimensional object (\emph{C}) where nodes, subgraphs, or
edges have specific values (see tbl.~\ref{tbl:methods}). For the
purposes of prediction, this low-dimensional object encodes feature
vectors for \emph{e.g.} the nodes. Embedding also allows to visualize
the structure in the data differently (\emph{D}), much like with a
principal component analysis.}\label{fig:embedding}
}
\end{figure}

@@

\begin{figure}
\hypertarget{fig:prediction}{%
\centering
\includegraphics{figures/conceptual_prediction.png}
\caption{From a low-dimensional feature vector (see
fig.~\ref{fig:embedding}), it is possible to develop predictive
approaches. Nodes in an ecological network are species, for which we can
leverage phylogenetic relatedness (\emph{e.g.} Strydom, Bouskila, et
al., 2021) or functional traits to fill the values of additional species
we would like to project in this space (here, I, J, K, and L) from the
embedding of known species (here, A, B, C, and D). Because embeddings
can be projected back to a graph, this allows to reconstruct a network
with these new species. This approach constitutes an instance of
transfer learning.}\label{fig:prediction}
}
\end{figure}

@@

The second series of problems are related to determining which area
should be used to infer the new metaweb in, as this determines the
species pool that must be used. Metawebs can be constructed by assigning
interactions in a list of species within geographic boundaries. The
upside of this approach is that information at the country level is
likely to be required for biodiversity assessments, as countries set
goals at the national level (Buxton et al., 2021), and as quantitative
instruments are designed to work at these scales (Turak et al., 2017);
specific strategies are often enacted at smaller scales, nested within a
specific country (Ray et al., 2021). But there is no guarantee that
these boundaries are meaningful. In fact, we do not have a satisfying
answer to the question of ``where does a food web stop?''; the most
promising solutions involve examining the spatial consistency of network
area relationships (Fortin et al., 2021; see \emph{e.g.} Galiana et al.,
2018, 2019, 2021), which is impossible in the absence of enough
information about the network itself. This suggests that inferred
metawebs should be further downscaled to allow for \emph{a posteriori}
analyses.

The final family of problems relates less to ecological concepts and
more to the praxis of ecological research. Operating under the context
of national divisions, in large parts of the world, reflects nothing
more than the legacy of settler colonialism. Indeed, the use of
ecological data is not an apolitical act (Nost \& Goldstein, 2021), as
data infrastructures tend to be designed to answer questions within
national boundaries, and their use often draws upon and reinforces
territorial statecraft; as per Machen \& Nost (2021), this is
particularly true when the output of ``algorithmic thinking''
(\emph{e.g.} relying on machine learning to generate knowledge) can be
re-used for governance (\emph{e.g.} enacting conservation decisions at
the national scale). We therefore recognize that predictive approaches
deployed at the continental scale, no matter their intent, originate in
the framework that contributed to the ongoing biodiversity crisis (Adam,
2014), reinforced environmental injustice (Choudry, 2013; Domínguez \&
Luoma, 2020), and \emph{e.g.} as on Turtle Island, should be replaced by
Indigenous principles of land management (Eichhorn et al., 2019; No'kmaq
et al., 2021). As we see artificial intelligence/machine learning being
increasingly mobilized to generate knowledge that is lacking for
conservation decisions (\emph{e.g.} Lamba et al., 2019; Mosebo Fernandes
et al., 2020), our discussion of these tools need to go beyond the
technical, and into the governance consequences they can have.

>>>>

Being able to infer \emph{potential} interactions could be the catalyst
for significant breakthroughs in our ability to start thinking about
species interaction networks over large spatial scales (Hortal et al.,
2015). Understanding species interactions holds enormous potential to
not only understand and more rapidly learn about species interactions
and metawebs, but also how changes in management of a single species may
impact non-target species. In a recent overview of the field of
ecological network prediction, Strydom, Catchen, et al. (2021)
identified two challenges of interest to the prediction of interactions
at large scales. First, there is a relative scarcity of relevant data in
most places globally -- paradoxically, this restricts our ability to
infer interactions to locations where inference is perhaps the least
required (and still leaves us unable to address regions without data);
second, accurate predictions often demand accurate predictors, and the
lack of methods that can leverage small amount of data is a serious
impediment to our global predictive ability. In most places, our most
reliable biodiversity knowledge is that of a species pool: through the
analysis of databases like GBIF or IUCN, it is possible to establish a
list of species in a region of interest; but establishing the
interactions between these species is difficult.

@@

\begin{figure}
\hypertarget{fig:embedding}{%
\centering
\includegraphics{figures/conceptual_embedding.png}
\caption{Overview of the embedding process. A network (\emph{A}),
represented as its adjacency matrix (\emph{B}), is converted into a
lower-dimensional object (\emph{C}) where nodes, subgraphs, or edges
have specific values (see tbl.~\ref{tbl:methods}). For the purposes of
prediction, this low-dimensional object encodes feature vectors for
\emph{e.g.} the nodes. Embedding also allows to visualize the structure
in the data differently (\emph{D}), much like with a principal component
analysis.}\label{fig:embedding}
}
\end{figure}

@@

\begin{figure}
\hypertarget{fig:prediction}{%
\centering
\includegraphics{figures/conceptual_prediction.png}
\caption{From a low-dimensional feature vector (see
fig.~\ref{fig:embedding}), it is possible to develop predictive
approaches. Nodes in an ecological network are species, for which we can
leverage phylogenetic relatedness (\emph{e.g.} Strydom, Bouskila, et
al., 2021) or functional traits to fill the values of additional species
we would like to project in this space (here, I, J, K, and L) from the
embedding of known species (here, A, B, C, and D). Because embeddings
can be projected back to a graph, this allows us to reconstruct a
network with these new species. This approach constitutes an instance of
transfer learning.}\label{fig:prediction}
}
\end{figure}

@@

The second series of problems relate to determining which area should be
used to infer the new metaweb in, as this determines the species pool
that must be used. Metawebs can be constructed by assigning interactions
in a list of species within geographic boundaries. The upside of this
approach is that information at the country level is likely to be
required for biodiversity assessments, as countries set goals at the
national level (Buxton et al., 2021), and as quantitative instruments
are designed to work at these scales (Turak et al., 2017); specific
strategies are often enacted at smaller scales, nested within a specific
country (Ray et al., 2021). But there is no guarantee that these
boundaries are meaningful. In fact, we do not have a satisfying answer
to the question of ``where does a food web stop?''; the most promising
solutions involve examining the spatial consistency of network area
relationships (Fortin et al., 2021; see \emph{e.g.} Galiana et al.,
2018, 2019, 2021), which is impossible in the absence of enough
information about the network itself. This suggests that inferred
metawebs should be further downscaled to allow for \emph{a posteriori}
analyses.

The final family of problems relates less to ecological concepts and
more to the praxis of ecological research. Operating under the context
of national divisions, in large parts of the world, reflects nothing
more than the legacy of settler colonialism. Indeed, the use of
ecological data is not an apolitical act (Nost \& Goldstein, 2021), as
data infrastructures tend to be designed to answer questions within
national boundaries, and their use often draws upon and reinforces
territorial statecraft. As per Machen \& Nost (2021), this is
particularly true when the output of ``algorithmic thinking''
(\emph{e.g.} relying on machine learning to generate knowledge) can be
re-used for governance (\emph{e.g.} enacting conservation decisions at
the national scale). We therefore recognize that predictive approaches
deployed at the continental scale, no matter their intent, originate in
the framework that contributed to the ongoing biodiversity crisis (Adam,
2014), reinforced environmental injustice (Choudry, 2013; Domínguez \&
Luoma, 2020), and \emph{e.g.} as on Turtle Island, should be replaced by
Indigenous principles of land management (Eichhorn et al., 2019; No'kmaq
et al., 2021). As we see artificial intelligence/machine learning being
increasingly mobilized to generate knowledge that is lacking for
conservation decisions (\emph{e.g.} Lamba et al., 2019; Mosebo Fernandes
et al., 2020), our discussion of these tools need to go beyond the
technical, and into the governance consequences they can have.
